---
title: "Statistical Inference Course Project Part 2"
author: "Janus"
date: "Monday, October 20, 2015"
output: pdf_document
---
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 

Load the ToothGrowth data and perform some basic exploratory data analyses 
Provide a basic summary of the data.
Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
State your conclusions and the assumptions needed for your conclusions. 

Some criteria that you will be evaluated on:

Question 1: Did you  perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?

Question 2: Did the student perform some relevant confidence intervals and/or tests?

Question 3: Were the results of the tests and/or intervals interpreted in the context of the problem correctly? 

Question 4: Did the student describe the assumptions needed for their conclusions?

Answer 1. Load the ToothGrowth data and perform some basic exploratory data analyses.

```{r}
library(ggplot2)
library(datasets)
x <- ToothGrowth
# convert dose to factor
x$dose <- as.factor(x$dose)

str(x)
```


```{r, echo=FALSE}
table(x$dose, x$supp)
```


```{r}
ggplot(data=x, aes(x=dose, y=len, fill=supp)) + geom_bar(stat="identity",) + facet_grid(. ~ supp) + xlab("Dose in miligrams") + ylab("Tooth length") + guides(fill=guide_legend(title="Supplement type"))
```


```{r}
ggplot(aes(x=dose, y=len), data=x) + geom_boxplot(aes(fill=dose)) + xlab("Dose in miligrams") + ylab("Tooth length") + guides(fill=guide_legend(title="Dose"))
```


```{r}
ggplot(aes(x=supp, y=len), data=x) + geom_boxplot(aes(fill=supp)) + xlab("Supplement type") + ylab("Tooth length") + guides(fill=guide_legend(title="Supplement type"))
```

Answer 2. Provide a basic summary of the data.

```{r}
summary(x)
```


Answer 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering).

```{r}
t.test(len ~ supp, data = x)
```

Null hypothesis can not be rejected as confindence intervals contain zero and p-value is 0.06.

Supplement types seems to have no impact on Tooth growth.

```{r}
# three groups as per dose level pairs
x.doses_0.5_1.0 <- subset (x, dose %in% c(0.5, 1.0)) 
x.doses_0.5_2.0 <- subset (x, dose %in% c(0.5, 2.0)) 
x.doses_1.0_2.0 <- subset (x, dose %in% c(1.0, 2.0)) 

# Check for dose levels (0.5, 1.0)
t.test(len ~ dose, data = x.doses_0.5_1.0)
```


```{r}
# Check for dose levels (0.5, 2.0)
t.test(len ~ dose, data = x.doses_0.5_2.0)
```


```{r}
# Check for dose levels (1.0, 2.0)
t.test(len ~ dose, data = x.doses_1.0_2.0)
```


The p-value is less than 0.05 and confindence intervals don't contian 0. The average toot length increases with an inceasing dose. The null hypothesis can be rejected.

Answer 4. State your conclusions and the assumptions needed for your conclusions.

Conclusions

Supplement type seem to have no impact on tooth growth.
Inreasing the dose level leads to increased tooth growth as well.

